Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-07-18 idp/documentation/paper.tex
continued working on usability section

		\subsection{Curve Design Process}

@@

			The influence which the design tool has on the success or failure of this whole process thus depends on the following key points.

			\begin{itemize}
				\item Can the language describe curves succinctly and uniquely?
				\item Is it easy for humans to `read' the language from curves?
				\item Is it easy for humans to `speak' the language to the software?
				\item Is it possible for software to build curves from descriptions in the language?
			\end{itemize}

			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software. Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.

@@

@@

@@

			% shortcomings of existing solutions (bézier, spiro (link to spiro promotion), NURBS)
			%   TODO: maybe correspond these shortcomings and their illustrations to the stuff in the evalution section
			%   TODO: research shortcomings that spiro splines do not solve, maybe we can have illustrations for that, too
			%   illustration: it's easy to create parabolas instead of circle arcs
			%   illustration: spirals need many control points and it's easy to include small bumps
			%   illustration: curvature continuity is not guaranteed

@@

			% with language being this important, it seems irresponsible to simply choose to use the mathematical parameters of the model as the language
			% moving from implementation detail to differential geometric specification
			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles

@@

>>>>

		\subsection{Curve Design Process}
		\label{section:CurveDesignProcess}

@@

			The influence which the design tool has on the success or failure of this whole process thus depends on the following criteria.

			\begin{enumerate}
				\item Can the language describe curves succinctly and uniquely?
				\item Is it easy for humans to `read' descriptions in the language from curves?
				\item Is it easy for humans to `speak' descriptions in the language to the software?
				\item Is it possible for software to build curves from descriptions in the language?
			\end{enumerate}

			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software. Ideally, the language should be very expressive, facilitating points 1 to 3, but it should still be possible to fulfill point 4 and derive mathematical curves from a description, a process which becomes increasingly difficult with more expressive description languages. Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.

@@

				While many people would not consider free-hand curve design a form of vector graphics curve design, it is well-suited to demonstrate how the model of the curve design process and the concept of a communication language that the designer uses to convey the curve to the software can be used to analyze the usability of curve design tools.

				Free-hand curves are created by the designer using some kind of pointing device (for instance, a mouse or a graphics tablet), to convey a large number of points that are on the curve to the software. For the simplicity of the argument, we will assume that the graphics software will simply connect these points using straight lines. Some vector graphics software will use Bézier curves instead of straight lines, but due to the large number of points involved as well as due to the low precision of free-hand input using a pointing device, this does not make much of a difference.

				Applying the criteria from section \ref{section:CurveDesignProcess} to free-hand curve design, it is immediately obvious that hundreds of points along the curve are not a succinct way of describing a curve, so criterion 1 is definitely not met very well. Criterion 2 on the other hand seems fine, since all the points that lie on the curve are immediately apparent when looking at it. Point 3 is looking promising aswell, since it's easy enough to input points into some vector graphics software, although this is severely hindered by the fact that there are so many points that need to be specified. Criterion 4 is very well met again, since the software has no problem constructing a polygon from a few hundred points.

				Note that this only applies if the software actually uses the points in the described fashion, that is, by building a curve that goes through all the points. If, on the other hand, the software is tasked with striking a balance between a curve that goes through as many points as possible while also remaining as smooth as possible, this has to be regarded as an entirely different communication language. In this case, the points the user creates using his pointing device are merely an intermediate representation, what is actually communicated to the software is the smooth motion the user makes while using the pointing device. The success of this approach depends largely on the specific algorithm the software uses to fit a smooth curve through the given points as well as the curve primitive that is eventually used to represent the final curve.

@@

				% TODO - illustrations
				% curves that are best described in terms of curvature are difficult to put down
				%   an excessive number of control points are needed to pin down curvature
				%     compromises smoothness
				%     makes the curve hard to change
				%     somewhat okay when tracing an existing image, since the shapes are already final and the source is likely smooth
				%   examples
				%     circle arcs (may have too much curvature in the middle, no constant curvature)
				%     spirals (fluttershy's mane)
				%     lugia's back line
				% changind direction at some node usually adversely affects neighboring spline segments
				%   changing the direction on the left side of a node with a long segment to the right screws up the right side
				%     bézier splines do not have much degrees of freedom left after direction continuity is ensured, even though curvature continuity may still be violated
				%     requires additional segments, tradeoff between smoothness and locality, not much tool support for retaining as much smoothness as possible while adding more freedom
				% curvature continuity is not guaranteed
				%   curves with lots of changes in curvature sign are affected the most
				%   examples
				%     lugia's belly

@@

				% TODO - illustrations
				% would in principle be well-suited for circle arcs and spirals and other shapes with linearly changing curvature, but curvature can only be controlled indirectly using points
				% tangent direction and curvature can only be controlled indirectly
				%   sometimes, many nodes need to be placed to convey the desired direction/curvature information
				%   specifying direction/curvature always comes with the unwanted side-effect of specifying additional points, possibly introducing unwanted bumpiness
				% sometimes, nice, smooth curves can be achieved using only few points
				%   it's not obvious where the points have to be specified such that one needs to specify as few as possible
				%   changing those 'minimal' curves is somewhat hampered by the fact that the purpose of each point may be non-obvious
				%     has bad locality since the points on the curve don't really specify points, but have an ulterior purpose
				% suffers from instability issues
				% does guarantee curvature continuity

@@

			% with language being this important, it seems irresponsible to have it determined by the mathematical parameters of the model as the language
			% moving from implementation detail to differential geometric specification
			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
            % TODO - diagram: process of going from user input and fairness to curve
			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles

@@

			% demonstrate euler spirals and bézier curves being embedded
			% revisit examples made when analyzing Bézier splines and Spiro splines
